---
title: "Ecoli mutants 2DTPP analysis - Version 2"
subtitle: "XXXX"
author: "Frank Stein"
date: "`r date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = F,warning=F,cache = T)
```

#Initializing R session
##Defining a working directory
We defined the working directory for the analysis. 
This directory should contain the output data from isobarQuant.
```{r eval=F}
setwd("~/home/path to the right file location")
```

##Defining some analysis functions

```{r}
add_col_to_df<-function(df,col_name,colvar_init=NA,Vector,
                        data_col="Experiment.ID",Vectornames=NULL,
                        create_column=T){
  if(create_column)df[,col_name]<-colvar_init
  df[,col_name]<-as.character(df[,col_name])
  if(length(Vectornames)==0)Vectornames<-Vector
  count=1
  for(i in Vector){
    print(i)
    x<-Vectornames[count][1]
    df[,col_name][grep(i,df[,data_col])]<-x
    count=count+1
  }
  df[,col_name]<-factor(df[,col_name],ordered=T,levels=c(colvar_init,Vector))
  return(df)
}
robustMean <- function(x){
  if(length(x) == 1){return(x)}
  else{
    return(smhuber(x)$mu)
  } 
}
estimate_mode <- function(x) {
  x<-x[!is.na(x)]
  d <- density(x)
  d$x[which.max(d$y)]
}
glog2 <- function(x) ((asinh(x)-log(2))/log(2))
```

##Loading packages

```{r}
library(plyr)
library(reshape)
library(vsn)
library(limma)
library(Biobase)
library(matrixStats)
library(fdrtool)
library(gridExtra)
library(tidyverse)
```

##Loading data and defining experimental conditions vector

```{r}
conditions<-read.csv("metadata.csv")
files<-unique(conditions$Path)
files<-files[file.exists(file.path("data",files))]
data<-NULL
for(i in 1:length(files)){
  file=files[i]
  x<-read.delim(file.path("./data/",file))
  x<-subset(x,!grepl("[K,k][R,r][T,t][0-9]",gene_name))
  x<-subset(x,!grepl("#+",protein_id))
  x<-subset(x,qupm>=2)
  x$Path<-file
  x<-x[,c("protein_id","description","gene_name","top3","Path","qupm",
          grep("signal_sum",names(x),value = T))]
  data<-rbind(data,x)
  rm(x)
}
rm(files,i,file)
names(data)
str(data)
```

#Restructuring data

```{r}
data2<-data
data2$description<-NULL
data2$protein_id<-NULL
data2<-unique(data2)
mdata<-melt(data2,id.vars = c("gene_name","top3","Path","qupm"),
            variable_name = "tmt.label")
rm(data2)
mdata$measurement<-gsub("_[0-9]+.+","",mdata$tmt.label)
mdata$tmt.label<-gsub("([a-z,A-Z]+_)+","",mdata$tmt.label)
mdata$tmt.label<-as.character(mdata$tmt.label)
mdata<-merge(mdata,conditions)
mdata$Experiment<-NULL;mdata$Path<-NULL
mdata<-subset(mdata,!is.na(gene_name))
mdata$average.top3<-"average.top3"
mdata$found<-"found.in.reps"
cdata<-cast(mdata,formula=gene_name~measurement+setup+Temperature+rep+tmt.label,
            value = "value",fun.aggregate = mean,na.rm=T)
cdata2<-cast(subset(mdata,measurement=="signal_sum"),
             formula=gene_name~average.top3,
             value = "top3",fun.aggregate = mean,na.rm=T)
cdata<-merge(cdata,cdata2)
cdata2<-cast(subset(mdata,measurement=="signal_sum"),formula=gene_name~found,
             value = "top3",fun.aggregate = length)
cdata2$found.in.reps<-cdata2$found.in.reps/11
cdata<-merge(cdata,cdata2)
#found.in.reps for each Temperature
cdata2<-cast(subset(mdata,measurement=="signal_sum"),
             formula=gene_name~found+Temperature,
             value = "top3",fun.aggregate = length)
cols<-names(cdata2)[names(cdata2)!="gene_name"]
for(i in cols){
  cdata2[,i]<-cdata2[,i]/11
}
cdata<-merge(cdata,cdata2)

mdata$max.qupm<-"max.qupm"
cdata2<-cast(subset(mdata,measurement=="signal_sum"),
             formula=gene_name~max.qupm,value = "qupm",
             fun.aggregate = max,na.rm=T)
cdata<-merge(cdata,cdata2)
#qupm for each Temperature
cdata2<-cast(subset(mdata,measurement=="signal_sum"),
             formula=gene_name~max.qupm+Temperature,value = "qupm",
             fun.aggregate = max,na.rm=T)
cdata<-merge(cdata,cdata2)
#min.qupm for each Temperature
mdata$max.qupm<-"min.qupm"
cdata2<-cast(subset(mdata,measurement=="signal_sum"),
             formula=gene_name~max.qupm+Temperature,value = "qupm",
             fun.aggregate = min,na.rm=T)
cdata<-merge(cdata,cdata2)
rm(cdata2)
```

#Filtering data
```{r}
dim(cdata)
cdata<-subset(cdata,max.qupm>=2&found.in.reps>=10&
                (found.in.reps_42>=2|found.in.reps_45.4>=2))
dim(cdata)
```

#Building expression set
For the analysis of proteomics data, many popular methods from the 
microarray-field are borrowed. In order to directly apply the methods, 
the proteomics data are stored in an "Expression Set" data type.

##Constructing assay data

```{r}
raw_data<-cdata
rownames(raw_data)<-raw_data$gene_name
raw_data$gene_name<-NULL
raw_data<-raw_data[,grep("signal_sum",names(raw_data),value=T)]
names(raw_data)<-gsub("signal_sum_","",names(raw_data))
raw_data<-as.data.frame(raw_data)
```

##Constructing metadata

```{r}
metadata<-data.frame(col.name=names(raw_data))
metadata<-add_col_to_df(metadata,data_col = "col.name",col_name = "tmt.label",
                        Vector = c("126","127L","127H","128L","128H","129L",
                                   "129H","130L","130H","131L","131H"))
metadata<-add_col_to_df(metadata,data_col = "col.name",col_name = "Temperature",
                        Vector = as.character(sort(unique(conditions$Temperature))))
metadata<-add_col_to_df(metadata,data_col = "col.name",col_name = "rep",
                        Vector = c("rep1","rep2","rep3","rep4","rep5"))
metadata<-merge(metadata,conditions,sort=F)
metadata$ID<-metadata$col.name
metadata$condition<-with(metadata,paste0(setup,"_",Temperature))
```

##Constructing feature data

```{r}
rownames(metadata)<-metadata$ID
colnames(raw_data)<-metadata$ID
```

##Transformation of raw signal_sum data

The log2 is computed from the raw signal intensities. Furthermore, 
infinite values are transformed into missing (NA) values.
```{r}
raw_data_m<-log2(as.matrix(raw_data))
raw_data_m[is.infinite((raw_data_m))]<-NA
raw_data_m[is.nan((raw_data_m))]<-NA
```

##Constructing the Expression set

```{r}
raw_dataE <- ExpressionSet(assayData = raw_data_m,
                           phenoData = AnnotatedDataFrame(metadata))
validObject(raw_dataE)
rm(raw_data_m,metadata,raw_data,mdata)
```

#Remove Batcheffects
Batcheffects are removed by fitting a linear model to the data that try to 
explain the replicates. The batch effect is than subtracted from the data 
applying the limma package.

```{r}
batchcleaned_raw_dataE<-raw_dataE
exprs(batchcleaned_raw_dataE)<-
  removeBatchEffect(exprs(batchcleaned_raw_dataE),
    batch=as.character(pData(batchcleaned_raw_dataE)$rep),
    design=model.matrix(~as.character(pData(batchcleaned_raw_dataE)$condition)))
```

#Normalization
The vsn package from Wolfgang Huber is used to apply a variance stabilization 
normalization method on the log2 raw data. Since it is a TPP experiment, 
each temperature is normalized separately. Therefore, the expression set 
has to be reconstructed again.
```{r}
norm_batchcleaned_raw_dataE<-batchcleaned_raw_dataE
for(ng in unique(conditions$Temperature)){
  print(ng)
  vsn.fit<-vsn2(2^exprs(norm_batchcleaned_raw_dataE[ ,
                                  norm_batchcleaned_raw_dataE$Temperature==ng]))
  test<-predict(vsn.fit,2^exprs(norm_batchcleaned_raw_dataE[ ,
                                  norm_batchcleaned_raw_dataE$Temperature==ng]))
  for(cn in colnames(test)){
    exprs(norm_batchcleaned_raw_dataE)[ ,cn]<-test[,cn]
  }
  ttest<-meanSdPlot(vsn.fit,plot = F)
  print(ttest$gg+ggtitle(ng))
  rm(vsn.fit)
}
rm(test,ng)
```

#Merge normalized data back into 'data'
```{r}
cd<-as.data.frame(2^exprs(batchcleaned_raw_dataE))
names(cd)<-paste0("batchcleaned_raw_signal_sum_",names(cd))
cd$gene_name<-rownames(cd)
cdata<-merge(cdata,cd)
rm(cd)
cd<-as.data.frame(2^exprs(norm_batchcleaned_raw_dataE))
names(cd)<-paste0("norm_signal_sum_",names(cd))
cd$gene_name<-rownames(cd)
cdata<-merge(cdata,cd)
rm(cd)
```

#Calculation of fold changes

```{r}
fc<-cdata[,c("gene_name",grep("norm_signal_sum",names(cdata),value=T),
             grep("max.qupm_",names(cdata),value=T))]
names(fc)[grepl("norm_signal_sum",names(fc))]<-gsub("_[0-9]+.$","",
                                                    names(fc)[grepl("norm_signal_sum",
                                                                    names(fc))])
names(fc)<-gsub("norm_signal_sum_","",names(fc))
cols<-ncol(fc)
initial.names<-names(fc)
temps<-as.character(unique(conditions$Temperature))
setups<-as.character(unique(conditions$Compound))
reps<-as.character(unique(conditions$rep))

for(temp in temps){
  for(setup in setups){
    for(rep in reps){
      name<-paste0(setup,".median_",temp,"_",rep,".setupr")
      if(paste0(setup,"_",temp,"_",rep)%in%initial.names){
        fc[,name]<-fc[,paste0(setup,"_",temp,"_",rep)]/
          apply(X=fc[,grep(paste0(temp,"_",rep),initial.names)],
                MARGIN = 1,FUN = median,na.rm=T)
        fc[,name][fc[,paste0("max.qupm_",temp)]<2]<-NA
      }
    }
  }
}
setups<-as.character(unique(conditions$setup))
for(setup in setups){
  for(temp in temps){
    for(rep in reps){
      if(paste0(setup,"_",temp,"_",rep)%in%initial.names){
        name<-paste0(setup,"_",temp,"_",rep,".42.tempr")
        fc[,name]<-fc[,paste0(setup,"_",temp,"_",rep)]/
          fc[,paste0(setup,"_42","_",rep)]
        fc[,name][fc[,paste0("max.qupm_",temp)]<2]<-NA
      }
    }
  }
}

fc<-fc[,-c(2:cols)]
rm(temps,setups,temp,setup,rep,reps,initial.names)
cdata<-merge(cdata,fc)
```

#Individual plots for each protein

```{r}
mfc<-melt(cdata[,c("gene_name","average.top3","max.qupm",
                   grep("^signal_sum",names(cdata),value=T),
                   grep("^batchcleaned_raw_signal_sum",names(cdata),value=T),
                   grep("norm_signal_sum",names(cdata),value=T),
                   grep("setupr$",names(cdata),value=T),
                   grep("tempr$",names(cdata),value=T))],
          id.vars = c("gene_name","average.top3","max.qupm"))
mfc<-add_col_to_df(mfc,col_name = "measurement",data_col = "variable",
                   Vector=c("signal_sum","batchcleaned_raw_signal_sum",
                            "norm_signal_sum","setupr","tempr"))
mfc$variable<-gsub(".+ignal_sum_","",mfc$variable)
mfc$variable<-gsub("_[0-9]+$","",mfc$variable)
mfc$variable<-gsub("_[0-9]+[H,L]$","",mfc$variable)
mfc$variable<-gsub(".setupr$","",mfc$variable)
mfc$variable<-gsub(".ctrl","",mfc$variable)
mfc$variable<-gsub(".42.tempr$","",mfc$variable)
mfc$rep<-mfc$variable
mfc$variable<-gsub("_rep[0-9]+$","",mfc$variable)
names(mfc)[grep("variable",names(mfc))]<-c("condition")
mfc$experiment<-mfc$condition
mfc$rep<-gsub(".+_","",mfc$rep)
mfc$Temperature<-gsub(".+_","",mfc$condition)
mfc$experiment<-gsub("_[0-9,.]+$","",mfc$experiment)
mfc<-na.omit(mfc)
mfc$measurement<-revalue(mfc$measurement,replace = c("setupr"="ctrl.ratio",
                                                     "tempr"="T42C.ratio"))
mfc$condition<-paste(gsub("_vs.+","",mfc$experiment),mfc$Temperature,sep = "_")
mfc$condition<-gsub(".median_","_",mfc$condition)
mfc$experiment<-gsub(".median","",mfc$experiment)

mfc2<-mfc%>%
  spread(key = measurement,value = value)

mfc3<-mfc2

conditions$condition<-with(conditions,paste0(setup,"_",Temperature))
mfc2<-merge(mfc2,conditions,all=T)
mfc2$temp<-as.numeric(as.character(mfc2$Temperature))
mfc2$Temperature<-factor(mfc2$Temperature)
write.csv(mfc2,file.path("..","full_datasets",
                              paste0(unique(conditions$Compound),
                                     "_full_data_long_V2.csv")))
```

#Normalization QC-plot

```{r fig.height=7}
ggplot(data=subset(mfc,measurement%in%c("signal_sum", 
                                        "batchcleaned_raw_signal_sum", 
                                        "norm_signal_sum")),
       aes(Temperature,log2(value),fill=experiment,
           group=paste(condition,rep)))+theme_bw(base_size=12)+
  geom_boxplot()+
  facet_wrap(~measurement,scale="free")
ggsave("QC_plot-normalization_V2.pdf",width=30,height=15)
ggplot(data=subset(mfc,!measurement%in%c("signal_sum", 
                                         "batchcleaned_raw_signal_sum",
                                         "norm_signal_sum")&!is.na(experiment)),
       aes(Temperature,log2(value),
           fill=experiment,
           group=paste(Temperature,experiment,rep)))+
  theme_bw(base_size=12)+geom_boxplot()+
  facet_wrap(~measurement,scale="free")
ggsave("QC_plot-normalization_fc_V2.pdf",width=15,height=15)
```

#LIMMA analysis

##Creating limma data set

```{r}
limma.score.data<-mfc2%>%
  filter(!is.na(ctrl.ratio))%>%
  group_by(gene_name,experiment,Temperature,rep)%>%
  summarise(ratio=median(ctrl.ratio))
limma.score.data.length<-mfc2%>%
  dplyr::filter(!is.na(ctrl.ratio),
                Temperature%in%sort(unique(mfc2$Temperature))[1:2])%>%
  group_by(gene_name,experiment,rep)%>%
  summarise(reps.abundance=length(ctrl.ratio))
limma.score.data<-left_join(limma.score.data,limma.score.data.length)
rm(limma.score.data.length)
limma.score.data.length<-mfc2%>%
  filter(!is.na(ctrl.ratio))%>%
  group_by(gene_name,experiment,rep)%>%
  summarise(No.of.Ts.with.2.reps=length(ctrl.ratio))#%>%
limma.score.data<-left_join(limma.score.data,limma.score.data.length)
rm(limma.score.data.length)
score.data.length<-mfc2%>%
  subset(!is.na(ctrl.ratio))%>%
  subset(Temperature%in%sort(unique(mfc2$Temperature))[1:2])%>%
  group_by(gene_name,experiment,Temperature)%>%
  summarise(rep=length(ctrl.ratio))%>%
  filter(rep>=2)%>%
  dplyr::select(gene_name,experiment)%>%
  group_by(gene_name,experiment)%>%
  summarise(Freq=length(experiment))%>%
  filter(Freq>=1)%>%
  dplyr::select(gene_name,experiment)%>%
  mutate(abundance.score.valid=T)
limma.score.data<-left_join(limma.score.data,score.data.length)
rm(score.data.length)
score.data.length<-mfc2%>%
  filter(!is.na(ctrl.ratio))%>%
  group_by(gene_name,experiment,Temperature)%>%
  summarise(rep=length(ctrl.ratio))%>%
  filter(rep>=2)%>%
  mutate(stability.score.valid=T)%>%
  dplyr::select(gene_name,experiment,Temperature,stability.score.valid)
limma.score.data<-left_join(limma.score.data,score.data.length)
rm(score.data.length)
abundance.score<-limma.score.data%>%
  subset(Temperature%in%sort(unique(limma.score.data$Temperature))[1:2])%>%
  subset(abundance.score.valid&stability.score.valid)%>%
  group_by(gene_name,experiment,rep)%>%
  summarise(abundance.score=mean(log2(ratio),na.rm=T))
limma.score.data<-left_join(limma.score.data,abundance.score)
limma.score.data<-limma.score.data%>%
  mutate(stability=log2(ratio)-abundance.score)
stability.score<-limma.score.data%>%
  subset(stability.score.valid)%>%
  group_by(gene_name,experiment,rep)%>%
  summarise(stability.score=sum(stability,na.rm=T))
limma.score.data<-left_join(limma.score.data,stability.score)
rm(abundance.score,stability.score)
limma.score.data<-limma.score.data%>%
  group_by(gene_name,experiment,rep)%>%
  dplyr::select(gene_name,experiment,reps.abundance,No.of.Ts.with.2.reps,
                abundance.score.valid,stability.score.valid,
                rep,abundance.score,stability.score)
limma.score.data<-unique(limma.score.data)
limma.score.data<-limma.score.data%>%
  mutate(id=paste0(experiment,"_",rep))%>%
  group_by(gene_name,id)%>%
  dplyr::select(gene_name,id,reps.abundance,No.of.Ts.with.2.reps,
                abundance.score.valid,stability.score.valid,
                abundance.score,stability.score)%>%
  filter(reps.abundance>=1,No.of.Ts.with.2.reps>=4)
write.csv(limma.score.data,file.path("..","raw_score_data",
                              paste0(unique(conditions$Compound),
                                     "_raw_score_data_V2.csv")))
```

##Constructing expression sets for abundance and stability data

```{r}
##abundance
abundance.data<-limma.score.data%>%
  group_by(gene_name)%>%
  dplyr::select(gene_name,id,abundance.score)%>%
  unique()%>%
  spread(id,abundance.score)
abundance.weights<-limma.score.data%>%
  group_by(gene_name)%>%
  dplyr::select(gene_name,id,reps.abundance)%>%
  unique()%>%
  spread(id,reps.abundance)
abundance.weights<-base::as.data.frame(abundance.weights)
rownames(abundance.weights)<-abundance.weights$gene_name
abundance.weights$gene_name<-NULL
abundance.weights<-as.matrix(abundance.weights)
abundance.data<-base::as.data.frame(abundance.data)
rownames(abundance.data)<-abundance.data$gene_name
abundance.data$gene_name<-NULL
abundance.metadata<-data.frame(ID=names(abundance.data))
abundance.metadata$condition<-gsub("_.+","",abundance.metadata$ID)
abundance.metadata$rep<-gsub(".+_","",abundance.metadata$ID)
rownames(abundance.metadata)<-abundance.metadata$ID
colnames(abundance.data)<-abundance.metadata$ID
abundance.data<-as.matrix(abundance.data)
abundance.dataE<- ExpressionSet(assayData = abundance.data,
                           phenoData = AnnotatedDataFrame(abundance.metadata))
validObject(abundance.dataE)
rm(abundance.data,abundance.metadata)

##stability
stability.data<-limma.score.data%>%
  group_by(gene_name)%>%
  dplyr::select(gene_name,id,stability.score)%>%
  unique()%>%
  spread(id,stability.score)
stability.weights<-limma.score.data%>%
  group_by(gene_name)%>%
  dplyr::select(gene_name,id,No.of.Ts.with.2.reps)%>%
  unique()%>%
  spread(id,No.of.Ts.with.2.reps)
stability.weights<-base::as.data.frame(stability.weights)
rownames(stability.weights)<-stability.weights$gene_name
stability.weights$gene_name<-NULL
stability.data<-base::as.data.frame(stability.data)
rownames(stability.data)<-stability.data$gene_name
stability.data$gene_name<-NULL
stability.metadata<-data.frame(ID=names(stability.data))
stability.metadata$condition<-gsub("_.+","",stability.metadata$ID)
stability.metadata$rep<-gsub(".+_","",stability.metadata$ID)
rownames(stability.metadata)<-stability.metadata$ID
colnames(stability.data)<-stability.metadata$ID
stability.data<-as.matrix(stability.data)
stability.dataE<- ExpressionSet(assayData = stability.data,
                           phenoData = AnnotatedDataFrame(stability.metadata))
validObject(stability.dataE)
rm(stability.data,stability.metadata)
```

##Limma evaluation of stability and abundance scores

```{r}
limma_results<-NULL
```

##Construction of design matrix and defining conditions that will be compared

```{r fig.height=7}
limma_data=abundance.dataE
condition<-factor(pData(limma_data)$condition,ordered=F)
replicate<-factor(pData(limma_data)$rep,ordered=F)
condition_d<-model.matrix(~1)
to_condition<-as.character(unique(conditions$Compound))
mut_comparison<-eBayes(lmFit(limma_data,weights = abundance.weights^2))
res <- limma::topTable(mut_comparison, sort.by = "t",  
                       coef = 1, number = Inf)
res$gene_name<-rownames(res)
names(res)[grep("P.Value",names(res))]<-"pvalue.limma"
names(res)[grep("adj.P.Val",names(res))]<-"fdr.limma"
res<-subset(res,!is.na(logFC))
fdr_res<-NULL
try(fdr_res <- fdrtool(res$t, plot =T, verbose = F))
if(length(fdr_res)>0){
  res$pvalue.fdrtool<-fdr_res$pval
  res$qval.fdrtool <- fdr_res$qval
  res$lfdr.fdrtool <- fdr_res$lfdr
}
res$comparison <- paste("abundance -",to_condition)
limma_results<-rbind(limma_results,res)

limma_data=stability.dataE
condition<-factor(pData(limma_data)$condition,ordered=F)
replicate<-factor(pData(limma_data)$rep,ordered=F)
condition_d<-model.matrix(~1)
to_condition<-as.character(unique(conditions$Compound))
mut_comparison<-eBayes(lmFit(limma_data,weights = stability.weights^2))
res <- limma::topTable(mut_comparison, sort.by = "t",  
                       coef = 1, number = Inf)
res$gene_name<-rownames(res)
names(res)[grep("P.Value",names(res))]<-"pvalue.limma"
names(res)[grep("adj.P.Val",names(res))]<-"fdr.limma"
res<-subset(res,!is.na(logFC))
fdr_res<-NULL
try(fdr_res <- fdrtool(res$t, plot =T, verbose = F))
if(length(fdr_res)>0){
  res$pvalue.fdrtool<-fdr_res$pval
  res$qval.fdrtool <- fdr_res$qval
  res$lfdr.fdrtool <- fdr_res$lfdr
}
res$comparison <- paste("stability -",to_condition)
limma_results<-rbind(limma_results,res)
```

##Limma and fdrtool statistics comparison

```{r fig.height=7}
ggplot(data=limma_results,aes(abs(t)))+
  geom_line(aes(y=fdr.limma,colour="limma - adj.P.Val"))+
  geom_line(aes(y=qval.fdrtool,colour="fdrtool - qval"))+
  geom_point(aes(y=fdr.limma,colour="limma - adj.P.Val"))+
  geom_point(aes(y=qval.fdrtool,colour="fdrtool - qval"))+
  facet_wrap(~comparison,scale="free_x")+ylab("fdr")+
  theme_bw(base_size=12)+
  scale_colour_brewer(palette="Set1")
ggsave("t_vs_fdr_limma_vs_fdrtool_V2.pdf",width=15,height=10)

ggplot(data=limma_results)+
  geom_histogram(aes(pvalue.limma,alpha=0.5,fill="limma"),bins=50)+
  facet_wrap(~comparison,scale="free_y")+
  geom_histogram(aes(pvalue.fdrtool,alpha=0.5,fill="fdrtool"),bins=50)+
  guides(alpha=F)+xlab("p-value")+
  theme_bw(base_size=12)+
  scale_fill_brewer(palette="Set1")
ggsave("p-value_histogram_limma_vs_fdrtool_V2.pdf",width=15,height=10)
```

##Hit annotation

```{r}
limma_results$hit_annotation_method<-NA
fdr_hit_threshold<-0.05
fdr_candidate_threshold=0.1
fc_hit_threshold<-1.5
fc_candidate_threshold<-1
limma_results$pvalue<-NA
limma_results$fdr<-NA
for(comp in unique(limma_results$comparison)){
  limma_hits<-
    nrow(limma_results[limma_results$comparison==comp&
                         limma_results$fdr.limma<=fdr_hit_threshold,])
  fdrtool_hits<-
    nrow(limma_results[limma_results$comparison==comp&
                         limma_results$qval.fdrtool<=fdr_hit_threshold,])
  if(limma_hits>=fdrtool_hits){
    limma_results$hit_annotation_method[limma_results$comparison==comp]<-"limma"
  }
  if(fdrtool_hits>limma_hits){
    limma_results$hit_annotation_method[limma_results$comparison==comp]<-"fdrtool"
  }
  rm(limma_hits,fdrtool_hits)
}
table(limma_results$hit_annotation_method)
limma_results$hit_annotation_method<-"fdrtool"

limma_results$pvalue[limma_results$hit_annotation_method=="limma"]<-
  limma_results$pvalue.limma[limma_results$hit_annotation_method=="limma"]
limma_results$fdr[limma_results$hit_annotation_method=="limma"]<-
  limma_results$fdr.limma[limma_results$hit_annotation_method=="limma"]

limma_results$pvalue[limma_results$hit_annotation_method=="fdrtool"]<-
  limma_results$pvalue.fdrtool[limma_results$hit_annotation_method=="fdrtool"]
limma_results$fdr[limma_results$hit_annotation_method=="fdrtool"]<-
  limma_results$qval.fdrtool[limma_results$hit_annotation_method=="fdrtool"]

limma_results$hit<-with(limma_results,
                        ifelse(fdr<=fdr_hit_threshold&abs(logFC)>=
                                 log2(fc_hit_threshold),T,F))
limma_results$hit_annotation<-with(limma_results,
                                   ifelse(fdr<=fdr_hit_threshold&abs(logFC)>=
                                            log2(fc_hit_threshold),"hit",
                                          ifelse(fdr<=fdr_candidate_threshold&abs(logFC)>=
                                                   log2(fc_candidate_threshold),
                                                 "candidate","no hit")))
limma_results$hit_annotation<-factor(limma_results$hit_annotation,ordered=T,
                                     levels=c("hit","candidate","no hit"))

with(limma_results,table(comparison,hit_annotation))
```

##Save limma results

```{r}
write.csv(limma_results,file.path("..","summarized_limma_results",
                              paste0(unique(conditions$Compound),
                                     "_Limma_results_V2.csv")))
write.csv(limma_results,"Limma_results_V2.csv",row.names=F)
```

##Merge limma significances with raw score data

```{r}
limma.cp.data<-limma.score.data%>%
  group_by(gene_name,id,reps.abundance,No.of.Ts.with.2.reps)%>%
  gather(key = score,value=score.value,-gene_name,-id,-reps.abundance,
         -No.of.Ts.with.2.reps,-stability.score.valid,-abundance.score.valid)
limma.cp.data$condition<-gsub("_.+","",limma.cp.data$id)
limma.cp.data$rep<-gsub(".+_","",limma.cp.data$id)
limma.sub<-limma_results[,c("gene_name","fdr","comparison")]
limma.sub$score<-gsub(" - .+",".score",limma.sub$comparison)
limma.sub$comparison<-NULL
limma.cp.data<-left_join(limma.cp.data,limma.sub)



limma.cp.data$fdr.star<-
  as.character(cut(as.numeric(as.character(limma.cp.data$fdr)),
                   breaks=c(0,0.001,0.01,0.05,1),
                   include.lowest = T))
limma.cp.data$fdr.star[is.na(limma.cp.data$fdr.star)]<-""
limma.cp.data$fdr.star<-revalue(limma.cp.data$fdr.star,replace = 
                                  c("(0.05,1]"="",
                                    "(0.01,0.05]"="*",
                                    "(0.001,0.01]"="**",
                                    "[0,0.001]"="***"))
```

##Volcano-Plot

```{r fig.height=7}
qplot(logFC,-log10(pvalue),data=limma_results,colour=hit_annotation)+
  geom_vline(aes(xintercept=0))+
  facet_wrap(~comparison+hit_annotation_method,scale="free_x")+
  geom_text(aes(label=gene_name),
            data=subset(limma_results,hit_annotation!="no hit"),
            vjust = 0, nudge_y = 0.1,size=2,check_overlap = F)+
  theme_bw(base_size=12)+
  scale_colour_brewer(palette="Set1")
ggsave("Volcano_plot_V2.pdf",width=15,height=10)
```

##Correlation of limma results

```{r fig.height=7}
cor.data<-limma_results[,c("gene_name","hit_annotation","logFC","comparison")]
cor.data$comparison<-gsub(" -.+","",cor.data$comparison)
cor.data$hit<-cor.data$hit_annotation
cor.data$hit_annotation<-NULL
cor.data<-unique(cor.data)
cor.data<-subset(cor.data,gene_name!="")
ccor.data<-cast(cor.data,formula=gene_name~comparison,value = "logFC")
ccor.data2<-cast(cor.data,formula=gene_name~comparison,value = "hit")
names(ccor.data2)[-1]<-paste0("hit_",names(ccor.data2)[-1])
ccor.data<-merge(ccor.data,ccor.data2)
ccor.data<-na.omit(ccor.data)
rm(cor.data,ccor.data2)
ccor.data$hit<-paste(ccor.data$hit_abundance,"abundance -",
                     ccor.data$hit_stability,"stability")
ccor.data$hit<-factor(ccor.data$hit,ordered=T,
                      levels=c("hit abundance - hit stability",
                               "hit abundance - candidate stability",
                               "hit abundance - no hit stability",
                               "candidate abundance - hit stability",
                               "no hit abundance - hit stability",
                               "candidate abundance - candidate stability",
                               "candidate abundance - no hit stability",
                               "no hit abundance - candidate stability",
                               "no hit abundance - no hit stability"))
ggplot(data=na.omit(ccor.data),aes(abundance,stability,colour=hit))+
  guides(alpha=F)+theme_bw(base_size=12)+
  geom_abline()+
  scale_color_brewer(palette="Set1")+
  geom_vline(aes(xintercept=0))+
  geom_hline(aes(yintercept=0))+
  geom_point(alpha=0.3)+
  geom_text(data=subset(ccor.data,abs(abundance)>=log2(1.5)|
                          abs(stability)>=log2(1.5)),aes(label=gene_name),
            vjust = 0, nudge_y = 0.1,size=2)+
  xlab("abundance score")+
  ylab("stability score")
ggsave("limma_correlation_plot_V2.pdf",width=20,height=20)
write.csv(ccor.data,"limma_correlation_data_V2.csv")
```

#Create Excel-Sheet for data browsing
```{r}
mfc3<-mfc2
mfc3<-subset(mfc3,signal_sum!=0)
#median ctrl.ratio
odata<-cast(subset(mfc3,as.character(Compound)==as.character(experiment)),
            formula=gene_name+Temperature~experiment+rep,
            value="ctrl.ratio")

#found in Temperatures
odata2<-cast(subset(mfc3,as.character(Compound)==as.character(experiment)),
             formula=gene_name~experiment,
             value="ctrl.ratio",fun.aggregate=length)
odata2$found.in.all.Ts<-odata2[,2]
odata2<-odata2[,c("gene_name","found.in.all.Ts")]
odata<-merge(odata,odata2,all.x=T)
rm(odata2)
#
odata2<-cast(subset(mfc3,as.character(Compound)==as.character(experiment)&
                      !is.na(ctrl.ratio)),
             formula=gene_name+Temperature~experiment,
             value="ctrl.ratio",fun.aggregate=length)
odata2$found.in.Ts<-odata2[,3]
odata2<-odata2[,c("gene_name","Temperature","found.in.Ts")]
odata<-merge(odata,odata2,all.x=T)
odata$found.in.Ts[is.na(odata$found.in.Ts)]<-0
rm(odata2)

modata<-melt(cdata[,c("gene_name","average.top3",
                      grep(".qupm_",names(cdata),value=T))],
             id.vars = c("gene_name","average.top3"))
modata$Temperature<-gsub(".+_","",modata$variable)
modata$variable<-gsub("_.+","",modata$variable)
modata<-cast(modata,formula=gene_name+average.top3+Temperature~variable)
odata<-merge(odata,modata,all.x=T)

odata$path<-file.path(".","plots_combined",
                      paste0(gsub("\\|","-",odata$gene_name),".pdf"))
des<-data[,c("gene_name","description")]
des<-subset(des,gene_name%in%unique(odata$gene_name))
des<-unique(des)
dupnames<-names(which(sort(table(des$gene_name),decreasing = T)>1))
des.dups<-subset(des,gene_name%in%dupnames)
des<-subset(des,!gene_name%in%dupnames)
des.dups.new<-NULL
for(gene in as.character(unique(des.dups$gene_name))){
  sub<-subset(des.dups,gene_name==gene)
  sub<-sub[which(nchar(sub$description)==max(nchar(sub$description)))[1],]
  sub<-data.frame(gene_name=sub$gene_name,description=sub$description)
  des.dups.new<-rbind(des.dups.new,sub)
  rm(sub)
}
des<-rbind(des,des.dups.new)
rm(des.dups.new,des.dups)
odata<-merge(odata,des)
rm(des)
odata<-odata[,c("gene_name","description","found.in.all.Ts", "found.in.Ts",
                "average.top3", "min.qupm","max.qupm","Temperature",
                grep("_rep.+$",names(odata),value=T),"path")]

write.csv(odata,"Final_results_V2.csv")

hd<-limma_results[,c("gene_name","comparison","hit_annotation")]
hd$score<-gsub(" -.+",".hit.annotation",hd$comparison)
hd$comparison<-NULL
hd<-cast(data=hd,formula=gene_name~score,value="hit_annotation")
rodata<-merge(odata,hd,all=T)
rodata<-rodata[order(rodata$gene_name,rodata$Temperature),]
rodata$gene_name2<-paste0("'",rodata$gene_name)
write.csv(rodata,"Final_results_classified_V2.csv")
rm(mfc3)


sodata<-unique(odata[,c("gene_name","description","found.in.all.Ts",
                        "average.top3","path")])
limma_spread<-limma_results[,c("gene_name","comparison","logFC","fdr",
                               "hit_annotation")]
limma_spread<-melt(limma_spread,id.vars = c("gene_name","comparison"))
limma_spread$score<-gsub(" - .+","",limma_spread$comparison)
limma_spread$comparison<-NULL
limma_spread<-cast(limma_spread,formula=gene_name~variable+score)
sodata<-merge(sodata,limma_spread,all=T)
write.csv(sodata,"Limma_results_classified_V2.csv")
```

#Save workspace
```{r}
save.image("Workspace_V2.RData")
```

#Save individual plots for each protein
```{r eval=F}
dir_save="plots_combined"
if(!dir.exists(file.path(getwd(),dir_save))){
  dir.create(file.path(getwd(),dir_save))
}
h=15
w=15
cols<-c("#e41a1c","#377eb8","#4daf4a","#984ea3","#ff7f00","#ffff33","#a65628",
        "#f781bf","#999999","#1b9e77","#d95f02","#7570b3","#e7298a","#66a61e",
        "#e6ab02","#a6761d")

limits.limma=quantile(limma.cp.data$score.value,probs = c(0.001,0.999),na.rm=T)
for(gene in as.character(unique(subset(cdata)$gene_name))){
  gene.name<-gsub("\\|","-",gene)
  f.name<-file.path(dir_save,paste0(gene.name,".pdf"))
  if(!file.exists(f.name)){
    pdf(width=w,height=h,file = f.name)
    sub<-subset(mfc2,gene_name==gene)
    sub$setup2<-as.character(sub$setup)
    sub$setup2[as.character(sub$setup)!=as.character(sub$Compound)]<-
      "average other mutants"
    sub$setup2<-paste(sub$setup2,sub$rep)
    average.top3=round(as.numeric(as.character(unique(sub$average.top3)[1])),digits = 2)
    max.qupm<-round(as.numeric(as.character(unique(sub$max.qupm)[1])))
    ptitle<-paste(gene,"-","average.top3:",average.top3,"-","max.qupm:",max.qupm)
    
    try(p1<-ggplot(data=subset(sub),aes(paste(rep,Temperature),setup))+
          geom_tile(aes(fill=log2(norm_signal_sum)))+
          scale_fill_gradientn(colours = c("#377eb8",
                                           "#984ea3",
                                           "#e41a1c",
                                           "#ff7f00",
                                           "#ffff33"))+
          xlab("condition")+
          theme_bw(base_size=12)+
          theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5)))
    try(p2<-ggplot(data=sub,aes(temp,log2(norm_signal_sum),fill=setup2,colour=setup2))+
          stat_summary(fun.data="mean_se",geom="smooth")+theme_bw(base_size=12)+ 
          theme(axis.text.x = element_text(angle = 90, hjust = 1))+
          scale_fill_manual(values = cols)+
          scale_colour_manual(values = cols)+
          stat_summary(fun.data="mean_se",geom="line",size=2))
    try(p3<-ggplot(data=subset(sub,experiment==as.character(unique(conditions$Compound)[1])),
                   aes(Temperature,paste(experiment,rep)))+
          geom_tile(aes(fill=log2(ctrl.ratio)))+
          scale_fill_gradient2(low="#2166ac",high="#b2182b",midpoint = 0,mid="#f7f7f7")+
          xlab("Temperature")+ylab("condition")+
          theme_bw(base_size=12)+
          theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5)))
    try(p4<-ggplot(data=sub,aes(temp,log2(T42C.ratio),fill=setup2,colour=setup2))+
          stat_summary(fun.data="mean_se",geom="smooth")+
          theme_bw(base_size=12)+ 
          theme(axis.text.x = element_text(angle = 90, hjust = 1))+
          scale_fill_manual(values = cols)+
          scale_colour_manual(values = cols)+
          stat_summary(fun.data="mean_se",geom="line",size=2))
    
    if(gene%in%limma.cp.data$gene_name){
      sub<-subset(limma.cp.data,gene_name==gene)
      new.limits<-limits.limma
      if(min(sub$score.value,na.rm=T)<new.limits[1]){
        new.limits[1]<-min(sub$score.value,na.rm=T)
      }
      if(max(sub$score.value,na.rm=T)>new.limits[2]){
        new.limits[2]<-max(sub$score.value,na.rm=T)
      }
      try(p5<-ggplot(data=subset(sub),aes(x=rep,score,fill=score.value))+
            geom_tile()+
            coord_polar()+
            scale_fill_gradient2(low="#2166ac",high="#b2182b",
                                 midpoint = 0,mid="#f7f7f7",
                                 limits=new.limits)+
            theme_bw(base_size=12)+
            xlab("replicate")+
            geom_text(aes(label=fdr.star)))
      try(p6<-ggplot(data=sub,aes(x=score,score.value,fill=score,group=score))+
            geom_boxplot()+
            geom_point(aes(shape=rep))+
            theme_bw(base_size=12)+
            scale_fill_brewer(palette="Set1")+
            # coord_cartesian(ylim = new.limits)+
            geom_text(aes(label=fdr.star),show.legend = FALSE)+
            geom_hline(yintercept=0)+
            theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5)))
      try(suppressWarnings( print(grid.arrange(p1, p2, p3, p4,p5,p6,ncol=2,top=ptitle))))
    }else{
      try(suppressWarnings( print(grid.arrange(p1, p2, p3, p4, ncol=2,top=ptitle))))
    }
    dev.off()
    rm(sub,min.p.value,new.limits,max.qupm,average.top3)
  }
}
```
